---
title: "PCoA_RA_plots"
author: "Carmen Saenz"
date: "5/16/2019"
output: html_document
---
<!-- Load required packages -->
```{r}
if("data.table" %in% rownames(installed.packages()) == FALSE) {install.packages("data.table")}
library(data.table)
if("tidyverse" %in% rownames(installed.packages()) == FALSE) {install.packages("tidyverse")}
library(tidyverse)
if("ggplot2" %in% rownames(installed.packages()) == FALSE) {install.packages("ggplot2")}
library(ggplot2)
if("ggrepel" %in% rownames(installed.packages()) == FALSE) {install.packages("ggrepel")}
library(ggrepel)
if("phyloseq" %in% rownames(installed.packages()) == FALSE) {install.packages("phyloseq")}
library(phyloseq)
if("reshape2" %in% rownames(installed.packages()) == FALSE) {install.packages("reshape2")}
library(reshape2)
library(tidyr)
library(vegan)
library(gtools)

set.seed(1234)
```

<!-- Set directories -->
```{r}
datapath <- "data/batch_culture/"
profile_param <- 'metaphlan' #'motus_rel' # 'metaphlan'
out_phyloseq = paste0(datapath, profile_param,'_phyloseq.rds')

```

<!-- function: PCoA plot -->
```{r}
plot_PCoA <- function(distance_lab, data_phyloseq, color, label){ #out_name,title_name,
  label2 <- noquote(label)
  #### PCoA
  ord<- ordinate(data_phyloseq, method = "PCoA", distance = distance_lab)
  PcoA_Sample_site_abundance <- plot_ordination(data_phyloseq, ord, color = color, shape=NULL, label=label) #type = type,
  PcoA_Sample_site_abundance <- PcoA_Sample_site_abundance +
    ggtitle(title_name, distance_lab)  +
    geom_point(size = 2.5) +
    theme_bw() +
    theme(legend.position = "top")
  PcoA_Sample_site_abundance$layers[[1]] <- NULL
  PcoA_Sample_site_abundance$layers[[2]] <- NULL
  PcoA_Sample_site_abundance$layers[[3]] <- NULL
  PcoA_Sample_site_abundance_2<- PcoA_Sample_site_abundance +
    geom_text_repel(aes(label = time), size = 3.0, segment.alpha = 0.5, max.overlaps = getOption("ggrepel.max.overlaps", default = 20)) +
    #geom_point(aes(fill=participant_ID2), colour="black", size = 2.5, shape = 21, alpha=0.7)
    geom_point(size = 2.1, shape = 16, alpha=1)
  PcoA_Sample_site_abundance_2$layers[[1]] <- NULL
  return(PcoA_Sample_site_abundance_2)
}

```
<!-- Load Phyloseq obj -->
```{r}
# Phyloseq object
if (profile_param %like% 'motus_rel') {
  profile_phyloseq_ra <- readRDS(out_phyloseq)
} else{
  profile_phyloseq <- readRDS(out_phyloseq)
  #### Transform counts to Relative Abundance
  profile_phyloseq_ra<-transform_sample_counts(profile_phyloseq, function(x){x/sum(x)})
} 
```

<!-- Phyloseq obj: metadata, count and taxa tables -->
```{r}
##### metadata:
sample_data_df <- data.frame(sample_data(profile_phyloseq_ra), stringsAsFactors = F)

### OTUs - db
taxa_table_df <- as.data.frame(unclass(tax_table(profile_phyloseq_ra)), stringsAsFactors = F)
taxa_table_df$taxa_ID <- rownames(taxa_table_df)
rownames(taxa_table_df) <- NULL

### Counts
otu_table_df <-  as.data.frame(unclass(otu_table(profile_phyloseq_ra)), stringsAsFactors = F)
otu_table_df$taxa_ID <- rownames(otu_table_df)
rownames(otu_table_df) <- NULL

otu_taxa_merge <- merge(otu_table_df, taxa_table_df, by = 'taxa_ID' , all.x = T)

# Filter by mOTUs present in the data
otu_taxa_filt_df = otu_taxa_merge[rowSums(otu_taxa_merge[, 2:ncol(otu_table_df)])>0,]
otu_taxa_filt_df = otu_taxa_filt_df %>% dplyr::select(1:ncol(otu_table_df))

otu_taxa_melt <- melt(otu_taxa_filt_df, id.vars=c("taxa_ID"))
otu_taxa_melt <- merge(taxa_table_df, otu_taxa_melt, by="taxa_ID")

otu_taxa_metadata =  merge(sample_data_df,otu_taxa_melt,  by.x = 'sample', by.y = 'variable')

otu_taxa_metadata$value <- as.numeric(otu_taxa_metadata$value)
```

# stats - C. difficile and C. scindens 
```{r}
# Cdiff and Cscind
summary(otu_taxa_metadata$value[which(otu_taxa_metadata$species == 'Clostridioides_difficile' & otu_taxa_metadata$condition == 'Cdiff')])
summary(otu_taxa_metadata$value[which(otu_taxa_metadata$species == 'Clostridioides_difficile' & otu_taxa_metadata$condition == 'Cdiff_Csind')])
summary(otu_taxa_metadata$value[which(otu_taxa_metadata$species == 'Clostridium_scindens' & otu_taxa_metadata$condition == 'Cdiff')])
summary(otu_taxa_metadata$value[which(otu_taxa_metadata$species == 'Clostridium_scindens' & otu_taxa_metadata$condition == 'Cdiff_Csind')])

summary(otu_taxa_metadata$value[which(otu_taxa_metadata$genus == 'Clostridium_scindens' & otu_taxa_metadata$condition == 'Cdiff')])
summary(otu_taxa_metadata$value[which(otu_taxa_metadata$genus == 'Clostridium_scindens' & otu_taxa_metadata$condition == 'Cdiff_Csind')])

otu_taxa_metadata_top15_cond_df <- data.frame(otu_taxa_metadata %>% 
                                           dplyr::group_by(genus, condition) %>% 
                                           dplyr::summarise(RA_count = sum(value)) %>% 
                                           dplyr::arrange(desc(RA_count))) 
summary(otu_taxa_metadata$value[which(otu_taxa_metadata$genus == 'Eisenbergiella' & otu_taxa_metadata$condition == 'Cdiff')])
summary(otu_taxa_metadata$value[which(otu_taxa_metadata$genus == 'Eisenbergiella' & otu_taxa_metadata$condition == 'Cdiff_Csind')])

```

# Plots
## Figure 1. C. scindens-mediated C. difficile growth reduction in an in vitro fermentor with two different culture setups. 
### C) Relative abundance of C. difficile and C. scindens when adding C. difficile alone (control) and C. difficile + C. scindens (treatment) to an existing fecal microbiota community during the treatment period of 42 hours in batch culture setup.
<!--Individual plots  -->
```{r}
# Prepare data
otu_taxa_metadata_top15 <- otu_taxa_metadata
otu_taxa_metadata_top15_sub <- otu_taxa_metadata_top15[otu_taxa_metadata_top15$species %in% c('Clostridioides_difficile', 'Clostridium_scindens'),]
unique(otu_taxa_metadata_top15_sub$species)

otu_taxa_metadata_top15_sub$species<- factor(otu_taxa_metadata_top15_sub$species, levels = c("Clostridioides_difficile", "Clostridium_scindens"))
otu_taxa_metadata_top15_sub$condition <- factor(otu_taxa_metadata_top15_sub$condition, levels = c('Cdiff', 'Cdiff_Csind'))

otu_taxa_metadata_top15_sub_plot <- otu_taxa_metadata_top15_sub
otu_taxa_metadata_top15_sub_plot$value[otu_taxa_metadata_top15_sub_plot$value == 0] <- NA
otu_taxa_metadata_top15_sub_plot <- otu_taxa_metadata_top15_sub_plot[!(otu_taxa_metadata_top15_sub_plot$condition=='Cdiff' & otu_taxa_metadata_top15_sub_plot$species=='Clostridium_scindens'),]

# Generate plot
colors_kit =c("Cdiff" = "#92392F", "Cdiff_Csind" = "#11698E") 

test_plot <- ggplot(data=otu_taxa_metadata_top15_sub_plot, aes(x=time, y=value, group = condition, color = condition)) +
  geom_point(alpha=0.5) + 
  #geom_bar(aes(y=RA_sum, fill = taxa_ID),stat="identity", alpha=.7) +
  theme_bw() + 
  theme(axis.text = element_text(size = 10),
        axis.text.x = element_text(color = "grey20", size = 10,angle = 60, face = "plain"),
        axis.text.y = element_text(color = "grey20", size = 10, angle = 0, hjust = 1, vjust = 0, face = "plain"),
        axis.title.x = element_text(color = "grey20", size = 12, angle = 0, hjust = 0.5, vjust = 1, face = "plain"),
        axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "plain"), 
        panel.grid.major.x = element_blank(), panel.grid.minor = element_blank(), legend.position="bottom") + 
  guides(color=guide_legend(nrow= 1))+ 
  xlab(NULL)+ylab("Relative abundance")+
  facet_grid(.~species)+
  scale_color_manual(values = colors_kit, name="Conditions")

print(test_plot +geom_smooth(method="loess",se=F, fill=NA)) 


```

<!-- Calculate fold change  -->
```{r}
## Cdiff
otu_taxa_metadata_top15_sub$value[(otu_taxa_metadata_top15_sub$condition=='Cdiff' & otu_taxa_metadata_top15_sub$time=='16'& otu_taxa_metadata_top15_sub$species=='Clostridioides_difficile')]/otu_taxa_metadata_top15_sub$value[(otu_taxa_metadata_top15_sub$condition=='Cdiff' & otu_taxa_metadata_top15_sub$time=='42'& otu_taxa_metadata_top15_sub$species=='Clostridioides_difficile')] #  0.4944606
## Cdiff_Csind
otu_taxa_metadata_top15_sub$value[(otu_taxa_metadata_top15_sub$condition=='Cdiff_Csind' & otu_taxa_metadata_top15_sub$time=='16'& otu_taxa_metadata_top15_sub$species=='Clostridioides_difficile')]/otu_taxa_metadata_top15_sub$value[(otu_taxa_metadata_top15_sub$condition=='Cdiff_Csind' & otu_taxa_metadata_top15_sub$time=='42'& otu_taxa_metadata_top15_sub$species=='Clostridioides_difficile')] # 3.645452

otu_taxa_metadata_top15_sub$value[(otu_taxa_metadata_top15_sub$condition=='Cdiff_Csind' & otu_taxa_metadata_top15_sub$time=='16'& otu_taxa_metadata_top15_sub$species=='Clostridium_scindens')]/otu_taxa_metadata_top15_sub$value[(otu_taxa_metadata_top15_sub$condition=='Cdiff_Csind' & otu_taxa_metadata_top15_sub$time=='42'& otu_taxa_metadata_top15_sub$species=='Clostridium_scindens')] # 2.808493

```

# Figure S2. Taxonomic profiles. 
## C) Relative abundance for the 15 most abundant genera across the samples in the batch culture setup using MetaPhlAn3 (Beghini et al., 2021).
<!-- Identified ASVs --> 
```{r}
# Prepare data
otu_taxa_metadata_top15_df <- data.frame(otu_taxa_metadata %>% 
                                           dplyr::group_by(genus) %>% 
                                           dplyr::summarise(RA_count = sum(value)) %>% 
                                           dplyr::arrange(desc(RA_count))) 

otu_taxa_metadata_top15_list = otu_taxa_metadata_top15_df$genus[1:15]
otu_taxa_metadata_top15_list <- otu_taxa_metadata_top15_list[otu_taxa_metadata_top15_list != "Unknown"][1:13]
otu_taxa_metadata_top15_list <- c(otu_taxa_metadata_top15_list, 'Clostridium_scindens', 'Clostridioides_difficile')

otu_taxa_metadata_top15 <- otu_taxa_metadata
otu_taxa_metadata_top15$genus[otu_taxa_metadata_top15$species=='Clostridioides_difficile'] <- 'Clostridioides_difficile'
otu_taxa_metadata_top15$genus[otu_taxa_metadata_top15$species=='Clostridium_scindens'] <- 'Clostridium_scindens'
otu_taxa_metadata_top15$genus[!otu_taxa_metadata_top15$genus %in% c(otu_taxa_metadata_top15_list, "Unknown")] <- 'Other'
otu_taxa_metadata_top15$genus[is.na(otu_taxa_metadata_top15$genus)] <- 'Other'

# Generate plot 
otu_taxa_metadata_top15_sum <- data.frame(otu_taxa_metadata_top15 %>% 
                                            dplyr::group_by(sample, condition, time, genus) %>% 
                                            dplyr::summarise(RA_count = sum(value)))

otu_taxa_metadata_top15_sum$genus<- factor(otu_taxa_metadata_top15_sum$genus, levels = rev(c(otu_taxa_metadata_top15_list, 'Other', 'Unknown')))

otu_taxa_metadata_top15_sum$condition <- factor(otu_taxa_metadata_top15_sum$condition, levels = c('Cdiff', 'Cdiff_Csind'))

otu_taxa_metadata_top15_sum$time <- as.character(otu_taxa_metadata_top15_sum$time)
day_order <- unique(otu_taxa_metadata_top15_sum$time)[order(as.numeric(unique(otu_taxa_metadata_top15_sum$time)))]
otu_taxa_metadata_top15_sum$time <- factor(otu_taxa_metadata_top15_sum$time, levels = day_order, ordered = TRUE)

colors_kit<-c('#D8DCDE', "#1C897E", "#CA5C2D", 
              '#B6D0E0',
              '#9D0208','#FFC87E','#F4A261','#E9C46A',
              '#A786C9','#D4C0E2','#975773','#6699FF','#000066',
              '#7AAFCA','#006699','#A9D181','#2F8475','#264445')

print(ggplot(data=otu_taxa_metadata_top15_sum, aes(x=time, group = genus)) +
        geom_bar(aes(y=RA_count, fill = genus),stat="identity", alpha=.9) +
        theme_minimal() + theme(axis.text = element_text(size = 8),panel.grid.minor = element_blank()) + labs(x = "Time (h)", y = "Relative abundance") +
        theme(title = element_text(size = 10),
              axis.text.x = element_text(color = "grey20", size = 10,angle = 0, face = "plain"),
              axis.text.y = element_text(color = "grey20", size = 10, angle = 0, hjust = 1, vjust = 0, face = "plain"),
              axis.title.x = element_text(color = "grey20", size = 12, angle = 0, hjust = 0.5, vjust = 1, face = "plain"),
              axis.title.y = element_text(color = "grey20", size = 12, angle = 90, hjust = .5, vjust = .5, face = "plain"), 
              panel.grid.major.x = element_blank(), panel.grid.minor = element_blank())+
        guides(fill=guide_legend(ncol= 1))+ 
        facet_grid(~condition, scales = "free_x")+
        scale_fill_manual(values = colors_kit, name="Genera"))

```

## D) Projection of the first two principal coordinates (explaining 66.4% and 28.3% of the total dissimilarity, respectively) based on Bray–Curtis dissimilarity between the microbiome composition in the control and treatment conditions of the batch culture setup.
<!--PCoA - Bray-Curtis -->
```{r}
# PCoA - Bray-Curtis
distance_lab = 'bray'
title_name <- c(paste0("PCoA - metaG"), "Bray-Curtis")
title_name_pval <- paste0("Bray-Curtis")
metadata_df <- as(sample_data(profile_phyloseq_ra), "data.frame")
if(length(unique(metadata_df$condition))>1){
  #**adonis/adonis2, Permutational Multivariate Analysis of Variance Using Distance Matrix**: ####
  library(vegan)
  otu_table_sort <- as.data.frame(unclass(t(otu_table(profile_phyloseq_ra))), stringsAsFactors = F)
  metadata_df <- data.frame(sample_data(profile_phyloseq_ra), stringsAsFactors = F)
  rownames(metadata_df) <- metadata_df$sample
  metadata <-metadata_df[match(rownames(otu_table_sort), rownames(metadata_df)),]
  adonis_bray_Status2<-adonis2(as.dist(vegdist(otu_table_sort, method="bray", na.rm = T)) ~ condition, data = metadata)
  r2_value <- format(round(adonis_bray_Status2$R2[1],3), nsmall = 3)
  p_value <- adonis_bray_Status2$`Pr(>F)`[1]
  
  title_name_pval <- paste0("Bray-Curtis; R2=",r2_value,"; pval=",p_value)
}

plot_PCoA_out <- plot_PCoA(distance_lab, profile_phyloseq_ra, "condition", "time")

manual_plot_colors =c("Cdiff" = "#92392F", "Cdiff_Csind" = "#11698E") 

print(plot_PCoA_out  + 
        scale_color_manual(values=manual_plot_colors, name='Condition') +
        #coord_fixed() +
        theme(legend.position="bottom")+
        ggtitle(title_name, title_name_pval))
# print(plot_PCoA_out  + 
#         facet_wrap(.~condition)+
#         scale_color_manual(values=manual_plot_colors, name='Condition') +
#         #coord_fixed() +
#         theme(legend.position="bottom")+
#         ggtitle(title_name, title_name_pval))
```

